Postural changes mediate children's visual access to social information

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd

### Face Detection Algorithms
The first algorithm was based on the work by @zhang2016 using multi-task cascaded convolutional neural networks (MTCNNs). The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present. We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet). This detector provided information about the presence of a face as well as its size and position.

The second algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc.) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose. The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose. For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose). We operationalized face detections as any frames containing a nose, as any half of a face necessarily contains a nose.  
<!-- we operationalized hand detections as any frames containing any of the 21 points detected on either the left or the right hand, and we operationalized wrist detections any frames containing either the left or the right wrist.   -->

### Detector evaluation
In order to evaluate the performance of these detectors, we constructed a "gold set" of frames by hand labeling both a sample of frames with a high density of faces (as reported by the detectors) and a random sample from the remaining frames. This was done so as to not bias our evaluation by the relatively rare appearance of faces in the dataset. A face was present in a frame if at least half of the face was showing. Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated and are reported in Table 2.
Both detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample. Due to the high performance of these two detectors we have included the results of both in the analyses to follow.

```{r, message=FALSE, warning=FALSE, results='asis'}
d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

@@

## couple of helper functions
evaluate <- function(a, b) {
  if (a == TRUE) {
    if (a == b) return ("TP") # was face, said face
    else return("FN") # was face, missed face
  }
  else {
    if (a == b) return("TN") # was not face, said not face
    else return("FP") # was not face, said face
  }
}

@@

## evaluate detectors
performance <- ground_truth_data  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong)   %>%
  mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face)) %>%
  group_by(detectorType,sample_type) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  group_by(detectorType,sample_type) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))
```

```{r xtable, results="asis"}
 tab2 <- xtable::xtable(performance, digits=3, caption = "Detector performance", type="latex", comment = F, table.placement = "H")
```

\begin{table}[ht]
\centering
\begin{tabular}{rllrrr}
  \hline
 & detectorType & sample\_type & p & r & f \\ 
  \hline
1 & MTCNN & high\_density & 0.841 & 1.000 & 0.914 \\ 
  2 & MTCNN & random & 0.947 & 0.750 & 0.837 \\ 
  3 & OpenPose & high\_density & 0.995 & 0.863 & 0.924 \\ 
  4 & OpenPose & random & 0.710 & 0.917 & 0.800 \\ 
  5 & ViolaJones & high\_density & 0.991 & 0.495 & 0.660 \\ 
  6 & ViolaJones & random & 0.389 & 0.292 & 0.333 \\ 
   \hline
\end{tabular}
\caption{Detector performance} 
\end{table}

@@

```{r echo=FALSE, include=FALSE}
# get rid of violajones
dlong <- dlong %>%
  filter(detectorType!="faceVJ")

>>>>

### Face Detection Algorithms
The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, @viola2004robust) applied to each individual frame. This detector provided information about the presence of a face as well as its size and position.

The second algorithm was based on the work by @zhang2016 using multi-task cascaded convolutional neural networks (MTCNNs). The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present. We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet). Like Viola-Jones, this detector provided information about the presence of a face as well as its size and position.

The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc.) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose. The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose. For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose). We operationalized face detections as any frames containing a nose, as any half of a face necessarily contains a nose.  
<!-- we operationalized hand detections as any frames containing any of the 21 points detected on either the left or the right hand, and we operationalized wrist detections any frames containing either the left or the right wrist.   -->

### Detector evaluation
In order to evaluate the performance of these detectors, we constructed a "gold set" of frames by hand labeling both a sample of frames with a high density of faces (as reported by the detectors) and a random sample from the remaining frames. This was done so as to not bias our evaluation by the relatively rare appearance of faces in the dataset. A face was present in a frame if at least half of the face was showing. Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated and are reported in Table 2.
Both OpenPose and MTCNN detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample. Due to the high performance of these two detectors we have included the results of both in the analyses to follow. The ViolaJones detector did not perform as well on either gold set, so the results associated with this detector were not included.

```{r, message=FALSE, warning=FALSE, results='asis'}
### Needs to be consolidated so that we're aren't rereading all of these files here...
# data_vj<- read_csv("../data/final_output/viola.csv") %>%
#   distinct(video, frame, .keep_all = TRUE)
# 
# data_mtcnn <- read_csv("../data/final_output/mtcnn3.csv") %>%
#   distinct(video, frame, .keep_all = TRUE)
# 
# data_openpose <- read_csv("../data/final_output/openpose_results_truncated_2.csv")%>%
#   distinct(name, frame, .keep_all = TRUE) %>%
#   mutate(frame = sprintf("%05d", as.numeric(frame) + 1)) %>%
#   mutate(is_face = ifelse(Nose_conf != 0, "True", "False")) %>%
#   select(group, name, frame, is_face) %>%
#   rename(video = name)

d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

@@

## couple of helper functions

evaluate <- function(a, b) {
  if (a == TRUE) {
    if (a == b) return ("TP") # was face, said face
    else return("FN") # was face, missed face
  }
  else {
    if (a == b) return("TN") # was not face, said not face
    else return("FP") # was not face, said face
  }
}

@@

performance <- ground_truth_data  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong)   %>%
  mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face)) %>%
  group_by(detectorType,sample_type) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  group_by(detectorType,sample_type) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))
```

```{r xtable, results="asis"}
 tab2 <- xtable::xtable(performance, digits=3, caption = "Detector performance")
```

<!-- \begin{table}[H] -->
<!-- \centering -->
<!-- \begin{tabular}{rrrr} -->
<!--   \hline -->
<!--   & P & R & F \\  -->
<!-- \hline -->
<!-- ViolaJones High Density Sample & 0.99 & 0.50 & 0.66 \\ -->
<!-- ViolaJones Random Sample & 0.39 & 0.29 & 0.33 \\ -->
<!-- OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\  -->
<!-- OpenPose Random Sample & 0.71 & 0.92 & 0.80 \\ -->
<!-- MTCNN High Density Sample & 0.84 & 1.00 & 0.91 \\  -->
<!-- MTCNN Random Sample & 0.95 & 0.75 & 0.84 \\  -->
<!--    \hline -->
<!-- \end{tabular} -->
<!-- \caption{Face detector performance}  -->
<!-- \end{table} -->

@@

```{r echo=FALSE, include=FALSE}
d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

## make long form version for plotting both face detections on same graph
dlong <- d %>%
  mutate(id = paste(subid,frame)) %>%
  gather(key = detectorType, value = detection, faceMT, faceOP) %>%
  mutate(detectorType=as.factor(detectorType))

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.tex

\begin{abstract}
The ability to process social cues--including eye gaze--is a critical
component of children's early language and cognitive development.
However, as children reach their first birthday, they begin to locomote
themselves, walking and exploring their visual environment in an
entirely new way. How do these postural and locomotive changes affect
children's access to the social information relevant for word-learning?
Here, we explore this question by using head-mounted cameras to record
infants' (8-16 months of age) egocentric visual perspective and modern
computer vision algorithms to detect the proportion of faces in infants'
environments. We find that infants' postural changes largely mediate
infants' access to faces, suggesting that these postural and locomotive
developments facilitate infants' emerging linguistic and social
capacities. Broadly, we suggest that the combined use of head-mounted
cameras and the application of novel deep learning algorithms is a
promising avenue for understanding the statistics of infants' visual and
linguistic experience.

@@

However, as children are learning their first words, their view of the
world is also changing radically (K. Adolph \& Berger, 2007). Infants'
motor abilities improve dramatically near the end of the first year of
life, allowing them to locomote independently and to determine what they
see. These motor changes have drastic consequences for what children
see: crawling and walking infants have radically different views of the
world. During spontaneous play in a laboratory playroom, toddlers are
more likely to look at the floor while crawling than while walking (J.
Franchak, Kretch, Soska, \& Adolph, 2011); walking vs.~crawling infants
tend to have full visual access to their environment and the people in
it (K. S. Kretch, Franchak, \& Adolph, 2014).

One possibility is that these motor improvements have strong
developmental cascades, impacting children's emerging social, cognitive,
and linguistic abilities (Iverson, 2010). Indeed, these postural changes
also impact how children interact with their mothers; walking infants
make different kinds of object-related bids for attention from their
mothers than crawling infants, and tend to hear more action directed
statements (e.g., ``open it'') (Karasik, Tamis-LeMonda, \& Adolph,
2014). More directly, in an observational study, Walle \& Campos (2014)
found that children who were able to walk had both higher receptive and
productive vocabularies. On their account, children's ability to stand
and independently locomote may fundamentally change their ability to
access the social information (e.g., faces, gaze) relative to children
who are still crawling and sitting. In other words, the ability to
access more detailed social information may allow infants to learn words
quicker and more efficiently, facilitating language growth.

Recent work has begun to use egocentric, head-mounted cameras to
document the visual experiences of infants and children --- which even
for walking children are radically different from the adult perspective
(and not easily predicted by our own adult intuitions (Clerkin, Hart,
Rehg, Yu, \& Smith, 2017; J. Franchak et al., 2011, Yoshida \& Smith
(2008))). Children's views tend to be restricted and to be dominated by
objects and hands much more than that of adults (Yoshida \& Smith,
2008), and both computational and empirical work suggest that this
restricted viewpoint may be more effective for learning objects and
their labels than the comparable adult perspective (Bambach, Crandall,
Smith, \& Yu, 2017; D. Yurovsky, Smith, \& Yu, in press). Further,
recent work also suggests dramatic changes in the child's perspective
over the first two years of life, as views transition from primarily
containing a close up view of faces to capturing views of hands paired
with the objects they are acting on (Fausey, Jayaraman, \& Smith, 2016).

Here, we directly ask whether the postural changes that infants
experience as they reach their first birthday change the availability of
the social information relevant for word learning. To do so, we recorded
the visual experience of a group of infants and children using
head-mounted cameras during a brief laboratory free-play session;
children's posture and orientation relative to their caregiver were also
recorded from a third-person perspective and hand-annotated. We
capitalize on recent improvements in face detection algorithms (K.
Zhang, Zhang, Li, \& Qiao, 2016, Cao, Simon, Wei, \& Sheikh (2017)) to
analyze the frequences of faces in the child's visual environment.

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
  & P & R & F \\ 
\hline
ViolaJones High Density Sample & 0.99 & 0.50 & 0.66 \\
ViolaJones Random Sample & 0.39 & 0.29 & 0.33 \\
OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\ 
OpenPose Random Sample & 0.71 & 0.92 & 0.80 \\
MTCNN High Density Sample & 0.84 & 1.00 & 0.91 \\ 
MTCNN Random Sample & 0.95 & 0.75 & 0.84 \\ 
   \hline
\end{tabular}
\caption{Face detector performance} 
\end{table}

>>>>

\begin{abstract}
The ability to process social cues--including eye gaze--is a critical
component of children's early language and cognitive development.
However, as children reach their first birthday, they begin to locomote
themselves, walking and exploring their visual environment in an
entirely new way. How do these postural and locomotive changes affect
children's access to the social information relevant for word-learning?
Here, we explore this question by using head-mounted cameras to record
infants' (8-16 months of age) egocentric visual perspective and use
state-of-the-art computer vision algorithms to detect the proportion of
faces in infants' environments. We find that infants' postural changes
largely mediate infants' access to faces, hinting that these postural
and locomotive developments facilitate infants' emerging linguistic and
social capacities. Broadly, we suggest that the combined use of
head-mounted cameras and the application of novel deep learning
algorithms is a promising avenue for understanding the statistics of
infants' visual and linguistic experience.

@@

However, as children are learning their first words, their view of the
world is also radically changing (K. Adolph \& Berger, 2007). Infants'
motor abilities improve dramatically near the end of the first year of
life, allowing them to locomote independently. These motor changes have
drastic consequences for what children see; crawling and walking infants
have radically different views of the world. For example, during
spontaneous play in a laboratory playroom, toddlers are more likely to
look at the floor while crawling than while walking (J. Franchak,
Kretch, Soska, \& Adolph, 2011); in general, walking infants tend to
have full visual access to their environment and the people in it, while
crawling infants do not (K. S. Kretch, Franchak, \& Adolph, 2014).

One possibility is that these motor improvements have strong
developmental cascades, impacting children's emerging social, cognitive,
and linguistic abilities (Iverson, 2010). Indeed, these postural changes
also impact how children interact with their mothers; walking
(vs.~crawling) infants make different kinds of object-related bids for
attention from their mothers and tend to hear more action directed
statements (e.g., ``open it'') (Karasik, Tamis-LeMonda, \& Adolph,
2014). Further, in an observational study, Walle \& Campos (2014) found
that children who were able to walk had both higher receptive and
productive vocabularies. On their account, children's ability to stand
and independently locomote may fundamentally change their ability to
access social information (e.g., faces, gaze) relative to children who
are still crawling and sitting. In other words, the ability to access
more detailed social information may allow infants to learn words
quicker and more efficiently, facilitating language growth.

Recent work has begun to use egocentric, head-mounted cameras to
document the visual experiences of infants and children --- which even
for walking children are radically different from the adult perspective
(and not easily predicted by our own adult intuitions) (Clerkin, Hart,
Rehg, Yu, \& Smith, 2017; J. Franchak et al., 2011; Yoshida \& Smith,
2008). Children's views tend to be restricted and to be dominated by
objects and hands much more than that of adults (Yoshida \& Smith,
2008), and both computational and empirical work suggest that this
restricted viewpoint may be more effective for learning objects and
their labels than the comparable adult perspective (Bambach, Crandall,
Smith, \& Yu, 2017; D. Yurovsky, Smith, \& Yu, in press). Further,
recent work also suggests dramatic changes in the child's perspective
over the first two years of life, as views transition from primarily
containing a close up view of faces to capturing views of hands paired
with the objects they are acting on (Fausey, Jayaraman, \& Smith, 2016).

Here, we directly ask whether the postural and locomotive developments
change the availability of the social information relevant for word
learning. To do so, we recorded the visual experience of a group of
infants and children using head-mounted cameras during a brief
laboratory free-play session; children's posture and orientation
relative to their caregiver were also recorded from a third-person
perspective and hand-annotated. We capitalize on recent improvements in
face detection algorithms (Cao, Simon, Wei, \& Sheikh, 2017; K. Zhang,
Zhang, Li, \& Qiao, 2016) to analyze the frequencies of faces in the
child's visual environment, both overall and relative to naming events
by their caregivers.

@@
